### 2022-02-16 paper/paper.tex
paper progress

%% build: latexmk -pdf -pvc paper.tex

\documentclass[acmsmall]{acmart}
%% \raggedbottom

%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a JOURNAL article.
\acmJournal{JACM}
\acmVolume{37}
\acmNumber{4}
\acmArticle{111}
\acmMonth{8}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%% --------------------------------------------------------------------------------

\usepackage{xcolor}
\usepackage{mathpartir}
\usepackage{todonotes}
\presetkeys{todonotes}{inline}{}
\usepackage{scalerel}
\usepackage{bm}
%% \usepackage{amssymb}

\newcommand{\mit}[1]{\mathit{#1}}
\newcommand{\msf}[1]{\mathsf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mbf}[1]{\mathbf{#1}}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ext}{\triangleright}
\newcommand{\Code}{\msf{Code}}
\newcommand{\El}{\msf{El}}
\newcommand{\lam}{\msf{lam}}
\newcommand{\app}{\msf{app}}
\newcommand{\NatElim}{\msf{NatElim}}
\newcommand{\y}{\msf{y}}

\newcommand{\Lift}{{\Uparrow}}
\newcommand{\spl}{{\sim}}
\newcommand{\qut}[1]{\langle #1\rangle}

\newcommand{\mbbc}{\mbb{C}}
\newcommand{\mbbo}{\mbb{O}}
\newcommand{\ob}{_\mbbo}

\renewcommand{\U}{\msf{U}}
\newcommand{\Con}{\msf{Con}}
\newcommand{\Sub}{\msf{Sub}}
\newcommand{\Ty}{\msf{Ty}}
\newcommand{\Tm}{\msf{Tm}}
\newcommand{\Cono}{\msf{Con}_{\mbbo}}
\newcommand{\Subo}{\msf{Sub}_{\mbbo}}
\newcommand{\Tyo}{\msf{Ty}_{\mbbo}}
\newcommand{\Tmo}{\msf{Tm}_{\mbbo}}
\newcommand{\hCon}{\wh{\msf{Con}}}
\newcommand{\hSub}{\wh{\msf{Sub}}}
\newcommand{\hTy}{\wh{\msf{Ty}}}
\newcommand{\hTm}{\wh{\msf{Tm}}}

\newcommand{\p}{\mathsf{p}}
\newcommand{\q}{\mathsf{q}}

\newcommand{\Bool}{\msf{Bool}}
\newcommand{\true}{\msf{true}}
\newcommand{\false}{\msf{false}}
\newcommand{\List}{\msf{List}}
\newcommand{\nil}{\msf{nil}}
\newcommand{\cons}{\msf{cons}}
\newcommand{\Nat}{\msf{Nat}}
\newcommand{\zero}{\msf{zero}}
\newcommand{\suc}{\msf{suc}}
\renewcommand{\tt}{\msf{tt}}
\newcommand{\fst}{\msf{fst}}
\newcommand{\snd}{\msf{snd}}
\newcommand{\mylet}{\msf{let}}
\newcommand{\emptycon}{\scaleobj{.75}\bullet}
\newcommand{\id}{\msf{id}}

\newcommand{\Set}{\mathsf{Set}}
\newcommand{\Rep}{\msf{Rep}}
\newcommand{\blank}{{\mathord{\hspace{1pt}\text{--}\hspace{1pt}}}}
\newcommand{\emb}[1]{\ulcorner#1\urcorner}

\newcommand{\Stage}{\msf{Stage}}
\newcommand{\hato}{\bm\hat{\mbbo}}
\newcommand{\ev}{\mbb{E}}
\newcommand{\re}{\mbb{R}}

\theoremstyle{remark}
\newtheorem{notation}{Notation}

\newcommand{\whset}{\wh{\Set}}
\newcommand{\rexti}{\re_{\ext_1}^{-1}}

%% --------------------------------------------------------------------------------

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

\title{Staged Compilation With Two-Level Type Theory}

%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{András Kovács}
\email{kovacsandras@inf.elte.hu}
\orcid{0000-0002-6375-9781}
\affiliation{%
  \institution{Eötvös Loránd University}
  \country{Hungary}
  \city{Budapest}
}

\begin{abstract}
  The aim of staged compilation is to enable metaprogramming in a way such that
  we have guarantees about the well-formedness of code output, and we can also
  mix together object-level and meta-level code in a concise and convenient
  manner. In this work, we observe that two-level type theory (2LTT), a system
  originally devised for the purpose of developing synthetic homotopy theory,
  also serves as a system for staged compilation with dependent types. 2LTT has
  numerous good properties for this use case: it has a concise specification,
  well-developed algebraic and categorical model theory, and it supports a wide
  range of language features both at the object and the meta level. First, we
  give an overview of 2LTT's features and applications in staging. Then, we
  present a staging algorithm and provide a proof of correctness. Our algorithm
  is ``staging-by-evaluation'', analogously to the technique of
  normalization-by-evaluation, in that staging is given by the evaluation of
  2LTT syntax in a semantic domain. The staging algorithm together with its
  correctness proof constitutes a proof of strong conservativity of 2LLT over
  the object theory. To our knowledge, this is the first system for staged
  compilation which supports full dependent types and unrestricted staging for
  types.
\end{abstract}

\begin{CCSXML}
\end{CCSXML}

\keywords{type theory, two-level type theory, staged compilation}
\maketitle

\section{Introduction}

The purpose of staged compilation is to write code-generating programs in a
safe, ergonomic and expressive way. It is always possible to do ad-hoc code
generation, by simply manipulating strings or syntax trees in a sufficiently
expressive programming language. However, these approaches tend to suffer from
verbosity, non-reusability and lack of safety. In staged compilation, there are
certain \emph{restrictions} on which metaprograms are expressible. Usually,
staged systems enforce typing disciple, prohibit arbitrary manipulation of
object-level scopes, and often they also prohibit accessing the internal
structure of object expressions. On the other hand, we get \emph{guarantees}
about the well-scoping or well-typing of the code output, and we are also able
to use concise syntax for embedding object-level code.

\emph{Two-level type theory}, or 2LTT in short, was described by Annekov,
Capriotti, Kraus and Sattler \cite{twolevel}, building on ideas from Vladimir
Voevodsky \cite{hts}. The motivation was to allow convenient metatheoretical
reasoning about a certain mathematical language (homotopy type theory), and to
enable concise and modular ways to extend the language with axioms.

It turns out that metamathematical convenience closely corresponds to to
metaprogramming convenience: 2LTT can be directly and effectively employed
in staged compilation. Moreover, semantic ideas underlying 2LTT are also
directly applicable to the theory of staging.

\subsection{Contributions}

\begin{itemize}
  \item In \ref{TODO} we present an informal syntax of two-level type theory, a
    dependent type theory with staging features. We look at basic use-cases
    involving inlining control, partial evaluation and fusion optimizations. We
    also describe feature variations, enabling applications in monomorphization
    and memory layout control.
  \item In \ref{TODO}, following \cite{twolevel}, we present a formal syntax of
    2LTT and the object theory (the target theory of code generation). We recall
    the standard presheaf model of 2LTT, which lies over the syntactic category
    of the object theory. We show that the evaluation of 2LTT syntax in the presheaf
    model yields a staging algorithm.
  \item In \ref{TODO} we show correctness of staging, consisting of
    \begin{itemize}
    \item \emph{Stability:} staging the output of staging has no action.
    \item \emph{Soundness:} the output of staging is convertible to the input.
    \item \emph{Completeness:} convertible programs produce convertible staging outputs.
    \end{itemize}
    Staging together with its correctness can be viewed as a \emph{strong
    conservativity} theorem of 2LTT over the object theory. This means that the
    possible object-level constructions in 2LTT are in bijection with the
    constructions in the object theory, and staging witnesses that meta-level
    constructions can be always computed away. This improves on the weak notion
    of conservativity shown in \cite{capriotti2017models} and
    \cite{twolevel}.
  \item To our knowledge, this is the first description of a language which
    supports staging in the presence of full-blown dependent types, with
    universes and large elimination. Moreover, we allow unrestricted staging
    for types, so that types can be computed by metaprograms at compile time.
\end{itemize}

\section{A Tour of Two-Level Type Theory}

In this section, we provide a short overview of 2LTT and its potential
applications in staging. We work in the informal syntax of a dependently typed
language which resembles Agda \cite{TODO}. We focus on examples and informal
explanations here; the formal details will be presented in Section \cite{TODO}.

\begin{notation}
We use the following notations throughout the paper. $(x : A) \to B$ denotes a
dependent function type, where $x$ may occur in $B$. We use $\lambda\,x.\,t$ for
abstraction. A $\Sigma$-type is written as $(x : A) \times B$, with pairing as
$(t,\,u)$, projections as $\fst$ and $\snd$, and we may use pattern matching
notation on pairs, e.g.\ as in $\lambda\,(x,\,y).\,t$. The unit type is $\top$
with element $\tt$. We will also use Agda-style notation for implicit arguments,
where $t : \{x : A\} \to B$ implies that the first argument to $t$ is inferred
by default, and we can override this by writing a $t \{u\}$ application. We may
also implicitly quantify over arguments (in the style of Idris and Haskell), for
example when declaring $\mit{id} : A \to A$ with the assumption that $A$ is
universally quantified.
\end{notation}

\subsection{Rules of 2LTT}

\paragraph{Universes}
We have universes $\U_{i,j}$, where $i \in \{0,1\}$, and $j \in \mbb{N}$.  The
$i$ index denotes stages, where $0$ is the runtime (object-level) stage, and $1$
is the compile time (meta-level) stage. The $j$ index denotes universe sizes in
the usual sense of type theory. We assume Russell-style universes, with
$\U_{i,j} : \U_{i, j+1}$. However, for the sake of brevity we will usually omit
the $j$ indices in this section, since sizing is orthogonal to our use-cases and
examples.
\begin{itemize}
\item $\U_0$ can be viewed as the \emph{universe of object-level or runtime types}.
    Each closed type $A : \U_0$ can be staged to an actual type in the object language
    (the language of the staging output).
  \item $\U_1$ can be viewed as the \emph{universe of meta-level or static types}. If we
    have $A : \U_1$, then $A$ is guaranteed to be only present at compile time,
    and will be staged away. Elements $a : A$ are likewise computed away.
\end{itemize}

\paragraph{Type formers} $\U_0$ and $\U_1$ may be closed under arbitrary type formers,
such as functions, $\Sigma$-types, identity types or inductive types in general.
However, all constructors and eliminators in type formers must stay at the same
stage. For example:
\begin{itemize}
  \item Function domain and codomain types must be at the same stage.
  \item If we have $\Nat_0 : \U_0$ for the runtime type of natural numbers,
        we can only map from it to a type in $\U_0$ by recursion or induction.
\end{itemize}
It is not required that we have the \emph{same} type formers at both stages. We
will discuss setups with different languages at different stages in Seection
\ref{todo}.

\paragraph{Moving between stages}
At this point, our system is rather limited, since there is no interaction
between the stages. We enable such interaction through the following operations.
\begin{itemize}
\item \emph{Lifting:} for $A : \U_0$, we have $\Lift A : \U_1$.  From the
  staging point of view, $\Lift A$ is the type of metaprograms which compute
  runtime expressions of type $A$.
\item \emph{Quoting:} for $A : \U_0$ and $t : A$, we have $\qut{t} :\,\Lift A$.
  A quoted term $\qut{t}$ represents the metaprogram which immediately yields
  $t$.
\item \emph{Splicing:} for $A : \U_0$ and $t :\,\Lift A$, we have $\spl t : A$.
  During staging, the metaprogram in the splice is executed, and the resulting
  expression is inserted into the output.
  \begin{notation} Splicing binds stronger than any operation, including function
    application. For instance, $\spl f\,x$ is parsed as $(\spl f)\,x$.
  \end{notation}

\item Quoting and splicing are definitional inverses, i.e.\ we have $\spl\qut{t} = t$ and
  $\qut{\spl t}=t$ as definitional equalities.
\end{itemize}
Note that none of these three operations can be expressed as functions, since
function types cannot cross between stages.

Informally, if we have a closed program $t : A$ with $A : \U_0$, \emph{staging}
means computing all metaprograms and recursively replacing all splices in $t$ and $A$
with the resulting runtime expressions. The rules of 2LTT ensure that this is possible,
and we always get a splice-free object program after staging.

\paragraph{Remark}
Why do we use the index $0$ for the runtime stage? The reason is that it is not
difficult to generalize 2LTT to multi-level type theory, by allowing to lift
types from $\U_i$ to $\U_{i+1}$. In the semantics, this can be modeled by
having a 2LTT whose object theory is once again a 2LTT, and doing this in an
iterated fashion. But there must be necessarily a bottom-most object theory;
hence our stage indexing scheme. For now though, we leave the multi-level
generalization to future work.

\begin{notation} We may disambiguate type formers at different stages by using $0$ or $1$
subscripts. For example, $\Nat_1 : \U_1$ is distinguished from $\Nat_0 : \U_0$,
and likewise we may write $\zero_0 : \Nat_0$ and so on. For function and
$\Sigma$ types, the stage is usually easy to infer, so we do not annotate
them. For example, the type $\Nat_0 \to \Nat_0$ must be at the runtime stage,
since the domain and codomain types are at that stage, and we know that the
function type former stays within a single stage. We may also omit stage annotations
from $\lambda$ and pairing.
\end{notation}

\subsection{Staged Programming in 2LTT}

In 2LTT, we may have several different polymorphic identity functions. First, we
consider the usual identity functions at each stage:
\begin{alignat*}{3}
  & \mit{id}_0 : (A : \U_0) \to A \to A\hspace{2em} && \mit{id}_1 : (A : \U_1) \to A \to A\\
  & \mit{id}_0 := \lambda\,A\,x.x       && \mit{id}_1 := \lambda\,A\,x.x
\end{alignat*}
An $\mit{id}_0$ application will simply appear in staging output as it is. In
contrast, $\mit{id}_1$ can be used as a compile-time evaluated function, because
the staging operations allow us to freely apply $\mit{id}_1$ to runtime
arguments. For example, $\mit{id}_1\,(\Lift\,\Bool_0)\,\qut{\true_0}$ has type
$\Lift \Bool_0$, therefore $\spl(\mit{id}_1\,(\Lift\,\Bool_0)\,\qut{\true_0})$
has type $\Bool_0$. We can stage this expression as follows:
\[
\spl(\mit{id}_1\,(\Lift\,\Bool_0)\,\qut{\true_0}) = \spl\qut{\true_0} = \true_0
\]
There is another identity function, which computes at compile time, but which
can be only used on runtime arguments:
\begin{alignat*}{3}
  & \mit{id_\Lift} : (A : \Lift\U_0) \to \Lift\,\spl A \to \Lift\,\spl A\\
  & \mit{id_\Lift} := \lambda\,A\,x.x
\end{alignat*}
Note that since $A : \Lift\U_0$, we have $\spl A : \U_0$, hence $\Lift\,\spl A
: \U_1$.  Also, $\Lift\U_0 : \U_1$, so all function domain and codomain types in
the type of $\mit{id_\Lift}$ are at the same stage. Now, we may write
$\spl(\mit{id_\Lift}\,\qut{\Bool_0}\,\qut{\true_0})$ for a term which is staged
to $\true_0$. In this specific case $\mit{id_\Lift}$ has no practical advantage
over $\mit{id}_1$, but in some cases we really have to quantify over
$\Lift\U_0$. This brings us to the next example.

Assume $\List_0 : \U_0 \to \U_0$ with $\nil_0 : (A : \U_0) \to \List_0\,A$,
$\cons_0 : (A : \U_0) \to A \to \List_0\,A$ and $\msf{foldr}_0 : (A\,B : \U_0)
\to (A \to B \to B) \to B \to \List_0\,A \to B$. We define a map function which
``inlines'' its function argument:
\begin{alignat*}{3}
  & \mit{map} : (A\,B : \Lift\U_0) \to (\Lift\,\spl A \to \Lift\,\spl B)
      \to \Lift(\List_0\,\spl A) \to \Lift(\List_0\,\spl B)\\
  & \mit{map} := \lambda\,A\,B\,f\,\mit{as}.\,
      \qut{\msf{foldr}_0\,
        \spl A\,\spl B\,
        (\lambda\,a\,\mit{bs}.\, \cons_0\,\spl B\,\spl(f\,\qut{a})\,\mit{bs})\,
        (\nil_0\,\spl B)\,
        \spl as
        }
\end{alignat*}
This $\mit{map}$ function can be defined with quantification over $\Lift \U_0$ but not
over $\U_1$, because $\List_0$ expects type parameters in $\U_0$, and there is
no generic way to convert from $\U_1$ to $\U_0$. Now, assuming
$\blank\!+_0\!\blank : \Nat_0 \to \Nat_0 \to \Nat_0$ and $\mit{ns} :
\List_0\,\Nat_0$, we have the following staging behavior:
\begin{alignat*}{3}
  & &&\spl(\mit{map}\,\qut{\Nat_0}\,\qut{\Nat_0}\,(\lambda\,n.\,\qut{\spl n +_0 10})\,\qut{\mit{ns}})\\
  & =\hspace{1em}&&\spl\qut{\msf{foldr}_0\,
        \spl \qut{\Nat_0}\,\spl \qut{\Nat_0}\,
        (\lambda\,a\,\mit{bs}.\, \cons_0\,\spl B\,\spl\qut{\spl\qut{a} +_0 10}\,\mit{bs})\,
        (\nil_0\,\spl \qut{\Nat_0})\,
        \spl\qut{\mit{ns}}}\\
  & =\hspace{1em} &&\msf{foldr_0}\,\Nat_0\,\Nat_0\,\,(\lambda\,a\,bs.\,a +_0 10)\,(\nil_0\,\Nat_0)\,\mit{ns}
\end{alignat*}

By using meta-level functions and lifted types, we already have control over
inlining. However, if we want to do more complicated meta-level computation, it
is convenient to use recursion or induction on meta-level type formers. A
classic example in staged compilation is the power function for natural numbers,
which evaluates the exponent at compile time. We assume the iterator function
$\msf{iter_1} : \{A : \U_1\} \to \Nat_1 \to (A \to A) \to A \to A$, and runtime
multiplication as $\blank\!*_0\!\blank$.
\begin{alignat*}{3}
  &\mit{exp} : \Nat_1 \to \Lift \Nat_0 \to \Lift \Nat_0 \\
  &\mit{exp} := \lambda\,x\,y.\,
  \msf{iter}_1\,x\,(\lambda\,n.\,\qut{\spl y *_0 \spl n})\,\qut{1}
\end{alignat*}
Now, $\spl(\mit{exp}\,3\,\qut{n})$ stages to $n *_0 n *_0 n *_0 1$ by the computation rules
of $\msf{iter_1}$ and the staging operations.

We can also stage \emph{types}. Below, we use iteration to compute the type of
vectors with static length, as a nested pair type.
\begin{alignat*}{3}
  &\mit{Vec} : \Nat_1 \to \Lift \U_0 \to \Lift \U_0\\
  &\mit{Vec} := \lambda\,n\,A.\,\msf{iter}_1\,n\,(\lambda\,B.\,\qut{\spl A \times \spl B})\,\qut{\top_0}
\end{alignat*}
With this definition, $\spl(\mit{Vec}\,3\,\qut{\Nat_0})$ stages to $\Nat_0
\times (\Nat_0 \times (\Nat_0 \times \top_0))$. Now, we can use \emph{induction}
on $\Nat_1$ to implement a map function. For readability, we use an Agda-style
pattern matching definition below (instead of the elimination principle).
\begin{alignat*}{6}
  &\hspace{-3em}\rlap{$\mit{map} : (n : \Nat_1) \to (\Lift\,\spl A \to \Lift\,\spl B) \to \Lift(\mit{Vec}\,n\,A) \to \Lift(\mit{Vec}\,n\,B)$}\\
  &\hspace{-3em}\mit{map}\,&&\zero_1      && f\,&&\mit{as} := \qut{\tt_0}\\
  &\hspace{-3em}\mit{map}\,&&(\suc_1\,n)\,&& f\,&&\mit{as} :=
     \qut{(\spl(f\,\qut{\fst_0\,\spl\mit{as}}),\,\mit{map}\,n\,f\,\qut{\snd_0\,\spl \mit{as}})}
\end{alignat*}
This definition inlines the mapping function for each projected element of the
vector. For instance, staging $\spl(\mit{map}\,2\,(\lambda\,n.\,\qut{\spl n +_0
  10})\,\qut{\mit{ns}})$ yields $(\fst_0\,\mit{ns} +_0
10,\,(\fst_0(\snd_0\,\mit{ns}) +_0 10,\,\tt_0))$. Sometimes, we do not want to
duplicate the code of the mapping function. In such cases, we can use
\emph{let-insertion} \cite{TODO}, a standard technique in staged compilation. If
we bind a runtime expression to a runtime variable, and only use that variable
in subsequent staging, only the variable itself can be duplicated. One solution
is to do an ad-hoc let-insertion:
\begin{alignat*}{6}
  &   && \mylet_0\,f := \lambda\,n.\, n +_0 10\,\,\msf{in}\,\,
         \spl(\mit{map}\,2\,(\lambda\,n.\,\qut{f\,\spl n})\,\qut{\mit{ns}}) \\
  & =\,\,&&  \mylet_0\,f := \lambda\,n.\, n +_0 10\,\,\msf{in}\,\,
          (f\,(\fst_0\,\mit{ns}),\,(f\,(\fst_0(\snd_0\,\mit{ns})),\,\tt_0))
\end{alignat*}
 Alternatively, we can define $\mit{map}$ so that it performs let-insertion, and
 we can switch between the two versions as needed.

More generally, we are free to use dependent types at the meta-level, so we can
reproduce more complicated staging examples. Any well-typed interpreter can be
rephrased as a \emph{partial evaluator}, as long as we have sufficient type
formers. For instance, we may write a partial evaluator for a simply typed
lambda calculus. We sketch the implementation in the following. First, we
inductively define contexts, types and terms:
\begin{alignat*}{6}
  & \Ty : \U_1  \hspace{2em} \Con : \U_1 \hspace{2em} \Tm : \Con \to \Ty \to \U_1
\end{alignat*}
Then we define the interpretation functions:
\begin{alignat*}{6}
  & \msf{EvalTy}  &&: \Ty \to \Lift \U_0 \\
  & \msf{EvalCon} &&: \Con \to \U_1 \\
  & \msf{EvalTm}  &&: \Tm\,\Gamma\,A \to \msf{EvalCon}\,\Gamma \to \Lift\,\spl(\msf{EvalTy}\,A)
\end{alignat*}
Types are necessarily computed to runtime types, e.g.\ an embedded
representation of the natural number type is evaluated to $\qut{\Nat_0}$.
Contexts are computed as follows:
\begin{alignat*}{4}
  &\msf{EvalCon}\,\,\msf{empty}                &&:= \top_1 \\
  &\msf{EvalCon}\,\,(\msf{extend}\,\Gamma\,A) &&:= \msf{EvalCon}\,\Gamma \times (\Lift \spl(\msf{EvalTy}\,A))
\end{alignat*}
This is an example for the usage of \emph{partially static
data}\cite{TODO}: semantic contexts are \emph{static} lists storing
\emph{runtime} expressions. This allows us to completely eliminate environment
lookups in the staging output: an embedded lambda expression is staged to the
corresponding lambda expression in the runtime language. This is similar to the
partial evaluator presented in Idris 1 \cite{TODO}. However, in contrast to
2LTT, Idris 1 does not provide a formal guarantee that partial evaluation does
not get stuck.

\subsection{Properties of Lifting \& Binding Time Improvements}

We describe more generally the action of $\Lift$ on type formers. First, $\Lift$
preserves negative type formers up to definitional isomorphism \cite{twolevel}:
\begin{alignat*}{5}
  \Lift((x : A) \to B\,x) &\simeq ((x : \Lift A) \to \Lift\,(B\,\spl x))\\
  \Lift ((x : A) \times B) &\simeq ((x : \Lift A) \times \Lift (B\,\spl x))\\
  \Lift \top_0 &\simeq \top_1
\end{alignat*}
For function types, the preservation maps are the following:
\begin{alignat*}{5}
  & \hspace{-6em}\rlap{$\mit{pres}_\to : \Lift((x : A) \to B\,x) \to ((x : \Lift A) \to \Lift\,(B\,\spl x))$}\\
  & \hspace{-6em}\mit{pres}_\to\,f     &&:= \lambda\,x.\,\,\qut{f\,\spl x}\\
  & \hspace{-6em}\mit{pres}_\to^{-1}\,f &&:= \qut{\lambda\,x.\,\,\spl(f\,\qut{x})}
\end{alignat*}
With this, we have that $\mit{pres}_\to\,(\mit{pres}_\to^{-1}\,f)$ is
definitionally equal to $f$, and also the other way around. Preservation maps
for $\Sigma$ and $\top$ work analogously.

By rewriting a 2LTT program left-to-right along preservation maps, we perform
what is termed \emph{binding time improvement} in the partial evaluation
literature \cite{TODO}. Note that the output of $\mit{pres}_{\to}$ uses a
meta-level $\lambda$, while going the other way we introduce a runtime
binder. Meta-level function and $\Sigma$ types support more computation during
staging, so in many cases it is beneficial to use the improved forms. In some
cases though we may want to use unimproved forms, to limit the size of generated
code. This is similar to what we have seen with let-insertion. For a minimal
example, consider the following unimproved version of $\mit{id}_\Lift$:
\begin{alignat*}{5}
  & \mit{id}_\Lift : (A : \Lift \U_0) \to \Lift(\spl A \to \spl A) \\
  & \mit{id}_\Lift := \lambda\,A.\,\,\qut{\lambda\,x.\,x}
\end{alignat*}
This can be used at the runtime stage as
$\spl(\mit{id}_\Lift\,\qut{\Bool_0})\,\true_0$, which is staged to
$(\lambda\,x.\,x)\,\true_0$.  This introduces a useless $\beta$-redex, so in
this case the improved version is clearly preferable.

For inductive types in general we only get oplax preservation. For example, we
have $\Bool_1 \to \Lift \Bool_0$, defined as
$\lambda\,b.\,\msf{if}\,b\,\msf{then}\,\qut{\true_0}\,\msf{then}\,\qut{\false_0}$.
In the staging literature, this is called ``serialization'' \cite{TODO}, or
``lifting'' in the context of Template Haskell \cite{TODO}. In the other
direction, we can only define constant functions from $\Lift \Bool_0$ to
$\Bool_1$.

The lack of elimination principles for $\Lift A$ means that we cannot inspect
the internal structure of expressions. This is called ``generativity'' in
staging \cite{TODO}. We will briefly discuss non-generative staging in Section
\ref{TODO}.

In particular, we have no serialization map from $\Nat_1 \to \Nat_1$ to
$\Lift(\Nat_0 \to \Nat_0)$. However, when $A : \U_1$ is \emph{finite}, and $B :
\U_1$ can be serialized, then $A \to B$ can be serialized, because it is
equivalent to a finite product. For instance, $\Bool_1 \to \Nat_1 \simeq \Nat_1
\times \Nat_1$.  In 2LTT, $A$ is called \emph{cofibrant} \cite{TODO}: this means
that for each $B$, $A \to \Lift\,B$ is equivalent to $\Lift C$ for some
$C$. This formalizes the so-called ``trick'' in partial evaluation, which
improves binding times by $\eta$-expanding functions out of finite sums
\cite{TODO}.

\subsubsection{Fusion}
Fusion optimizations can be viewed as binding time improvement techniques for
general inductive types. The basic idea is that by lambda-encoding an inductive
type, it is brought to a form which can be binding-time improved. For instance,
consider foldr-build fusion for lists, which is employed in GHC Haskell
\cite{TODO}. Starting from $\Lift (\List_0\,A)$, we use Böhm-Berarducci
encoding \cite{TODO} under the lifting to get
\[ \Lift((L : \U_0) \to (A \to L \to L) \to L \to L) \]
which is isomorphic to
\[ (L : \Lift\,U_0) \to (\Lift\,A \to \Lift\,\spl L \to \Lift\,\spl L)
    \to \Lift\,\spl L \to \Lift\,\spl L. \]
Alternatively, for \emph{stream fusion}, we embed $\List\,A$ into the
coinductive colists (i.e.\ the possibly infinite lists), and use a terminal
lambda-encoding. The embedding into the ``larger'' structure enables some staged
optimizations which are otherwise not possible, such as fusion for the
$\msf{zip}$ function \cite{TODO}. However, the price we pay is that converting
back to lists from colists is not necessarily total.

We do not detail the implementation of fusion in 2LTT here. In short, 2LTT is a
natural setting for a wide range of fusion setups. A major advantage of fusion
in 2LTT is the formal guarantee of staging, in contrast to implementations where
compile-time computation relies on ad-hoc user annotations and general-purpose
optimization passes. For instance, fusion in GHC relies on rewrite rules and
inlining annotations which have to be carefully tuned and ordered, and it is
quite possible to get pessimized code via failed fusion.

\subsubsection{Inferring staging operations}
During bidirectional elaboration \cite{TODO}, we can use the preservation
isomorphisms and the quote-splice isomorphism as a coercive subtyping
system. When elaboration needs to compare an inferred and an expected type, it
may insert transports along isomorphisms. We implemented this feature in our
prototype. It additionally supports Agda-style implicit arguments and pattern
unification, so it can elaborate the following definition:
\begin{alignat*}{3}
  & \mit{map} : \{A\,B : \Lift\U_0\} \to (\Lift A \to \Lift B)
      \to \Lift(\List_0\,A) \to \Lift(\List_0\,B)\\
  & \mit{map} := \lambda\,f\,\mit{as}.\,
      \msf{foldr}_0
        (\lambda\,a\,\mit{bs}.\, \cons_0\,(f\,a)\,\mit{bs})\,
        \nil_0\,
        as
\end{alignat*}
We may go a bit further, and also add the coercive subtyping rule $\U_0 \leq
\U_1$, witnessed by $\Lift$. Then, the type of $\mit{map}$ can be written as
$\{A\,B : \Lift\U_0\} \to (A \to B) \to \List_0\,A \to \List_0\,B$. However,
here the elaborator has to make a choice, whether to elaborate to improved or
unimproved types. In this case, the fully unimproved type would be
\[ \{A\,B : \Lift\U_0\} \to \Lift((\spl A \to \spl B) \to \List_0\,\spl A \to \List_0\,\spl B). \]
It seems to the author of this paper that improved types are a sensible default,
and we can insert explicit lifting when we want to opt for unimproved
types. This is also available in our prototype.

\subsection{Variations of Object-Level Languages}

%% In principle, 2LTT supports any sensible type-theoretic feature on the meta
%% level. On the object level though, we have yet more freedom in choosing type
%% formers and even the equational theory of terms. We may choose to omit
%% $\beta\eta$-rules from the object language.

%% \todo{merge closure typing into rep poly?}
%% \todo{range of object languages?}
%% \todo{effectful, etc object languages?}

In the following, we consider variations on object-level languages, with a focus
on applications in downstream compilation after staging. Adding restrictions or
more distinctions to the object language can make it easier to optimize and
compile it.

%% Simpler object languages are easier to optimize and compile, but they are often
%% more tedious to directly program in. 2LTT can be used as a principled way to
%% extended languages with abstraction facilities, through metaprogramming. We
%% look at several setups with restricted object languages.

\subsubsection{Monomorphization}

In this case, the object language is simply typed, so every type is known
statically. This makes it easy to assign different memory layouts to different
types, and generate code accordingly for each type. Moving to 2LTT, we still
want to abstract over runtime types at compile time, so we use the following
setup.
\begin{itemize}
\item We have a \emph{jugdment}, written as $A\,\msf{type_0}$, for well-formed
  runtime types. Runtime types may be closed under simple type formers.
\item We have a type $\Ty_0 : \U_1$ in lieu of the previous $\Lift \U_0$.
\item For each $A\,\msf{type}_0$, we have $\Lift A : \Ty_0$.
\item We have quoting and splicing for types and terms. For types, we send $A\,
  \msf{type}_0$ to $\qut{A} : \Ty_0$. For terms, we send $t : A$ to $\qut{t} :
  \Lift A$.
\end{itemize}
Despite the restriction to simple types at runtime, we can still write arbitrary
higher-rank polymorphic functions in this setup, such as a function with type
$((A : \Ty_0) \to \Lift\,\spl A \to \Lift \spl A) \to \Lift\,\Bool_0$. This
function can be only applied to statically known arguments, so the polymorphism
can be staged away. The main restriction that programmers have to keep in mind
is that polymorphic functions cannot be stored inside runtime data types.

This setup is fairly similar to the well-known monomorphization models in the
C++ and Rust programming languages. The evident difference is that in 2LTT we
can use a full type theory at compile time, as opposed to languages which are
less expressive and less principled. Additionally, we will see in Section
\ref{TODO} that this setup is compatible with an \emph{induction principle} for
$\Ty_0$, so that we can analyze the structure of runtime types.

\subsubsection{Memory representation polymorphism}

This refines monomorphization, in that types are not directly identified with
memory representations, but instead representations are internalized in 2LTT as
a meta-level type, and runtime types are indexed over representations.
\begin{itemize}
\item We have $\Rep : \U_1$ as the type of memory representations. We have
  considerable freedom in the specification of $\Rep$. A simple setup may
  distinguish references from unboxed products, i.e.\ we have $\msf{Ref} : \Rep$
  and $\msf{Prod} : \Rep \to \Rep \to \Rep$, and additionally we may assume
  any desired primitive machine representation as a $\Rep$.
\item We have Russell-style $\U_{0,j} : \Rep \to \U_{0, j+1}\,r$, where $r$ is
  some chosen runtime representation for types; often this would mark types as
  erased at runtime. We leave the meta-level $\U_{1,j}$ hierarchy unchanged.
\item We may introduce unboxed $\Sigma$ types and primitive machine types in the
  runtime language. For $A : \U_{0}\,r$ and $B : A \to \U_{0}\,r'$, we may have
  $(x : A) \times B\,x : \U_{0}\,(\msf{Prod\,r\,r'})$. Thus, we have
  type dependency, but we do not have dependency in memory representations.
\end{itemize}
Since $\Rep$ is meta-level, there is no way to abstract over it at runtime, and
during staging all $\Rep$ indices are computed to concrete canonical
representations. This is a way to reconcile dependent types with control over
memory layouts.  The unboxed flavor of $\Sigma$ ends up with a statically known
flat memory representation, computed from the representations of the fields.

In principle, it should be possible to have \emph{dependent memory layouts};
this would be a more advanced variation on type-passing polymorphism
\cite{TODO}, where layouts of objects may depend on runtime data. For example,
length-prefixed flat arrays would not be a primitive notion, but simply defined
as $(n : \Nat) \times \msf{Vec}\,n\,A$. However, it seems to be highly
challenging to implement runtime systems and precise garbage collection in this
setting. See \texttt{sixten} \cite{TODO} for an experimental implementation of
dependent memory layouts.

\section{Formal 2LTT}

In this section we switch to a formal description of 2LTT, which we will use in
the subsequent sections to define the staging algorithm and prove its
correctness. First we describe the metatheory that we work in.

\subsection{Metatheory}
TODO

\subsection{Models and Syntax of 2LTT}
We use an algebraic specification for models, and specify the syntax as the
initial model. We only handle well-formed syntactic objects and only define
operations which respect syntactic definitional equality; in fact, we identify
definitional equality with metatheoretic equality. First, we define the
structural scaffolding of 2LTT without type formers.

\begin{definition} A model of \textbf{basic 2LTT} consists of the following.
\begin{itemize}
\item
  A category $\mbbc$ with a terminal object. We denote the set of objects as
  $\Con_{\mbbc} : \Set$ and use capital Greek letters starting from $\Gamma$ to
  refer to objects. The set of morphisms is $\Sub_{\mbbc} : \Con_{\mbbc} \to
  \Con_{\mbbc} \to \Set$, and we use $\sigma$, $\delta$ and so on to refer to
  morphisms. The terminal object is written as $\emptycon$ with unique morphism
  $\epsilon : \Sub_\mbbc\,\Gamma\,\emptycon$. We omit the $\mbbc$ subscript if
  it is clear from context.
\item
  For each $i \in \{0,1\}$ and $j \in \mbb{N}$, we have $\Ty_{i,j} : \Con
  \to \Set$ and $\Tm_{i,j} : (\Gamma : \Con) \to \Ty_{i,j}\,\Gamma \to \Set$,
  where $\Ty$ is a presheaf over $\mbbc$ and $\Tm_{i,j}$ is a presheaf over the
  category of elements of $\Ty_{i,j}$. This means that both types ($\Ty$) and
  terms ($\Tm$) can be substituted, and substitution has functorial action. We
  use $A$, $B$, $C$ to refer to types and $t$, $u$, $v$ to refer to terms, and
  use $A[\sigma]$ and $t[\sigma]$ for substituting types and
  terms. Additionally, for each $\Gamma : \Con$ and $A : \Ty_{i,j}\,\Gamma$, we
  have the extended object $\Gamma \ext A : \Con$ such that there is a natural
  isomorphism $\Sub\,\Gamma\,(\Delta\ext A) \simeq ((\sigma :
  \Sub\,\Gamma\,\Delta) \times \Tm_{i,j}\,\Gamma\,(A[\sigma]))$.
\item
  For each $j$ we have a \emph{lifting structure}, consisting of a natural
  transformation $\Lift : \Ty_{0,j}\,\Gamma \to \Ty_{1,j}\,\Gamma$, and an
  invertible natural transformation $\qut{\blank} : \Tm_{0,j}\,\Gamma\,A \to
  \Tm_{1,j}\,\Gamma\,(\Lift A)$, with inverse $\spl\blank$.
\end{itemize}
\end{definition}

In short, for each $i$ and $j$ we have a family structure in the sense of
categories-with-families \cite{TODO}, such that there is a family morphism from
each $(\Ty_{0,j},\,\Tm_{0,j})$ to $(\Ty_{1,j},\,\Tm_{1,j})$ in the sense of
\cite{TODO}.

The following notions are derivable in any model:
\begin{itemize}
\item
  By moving left-to-right along $\Sub\,\Gamma\,(\Delta\ext A) \simeq ((\sigma :
  \Sub\,\Gamma\,\Delta) \times \Tm_{i,j}\,\Gamma\,(A[\sigma]))$,
  and starting from the identity morphism $\id : \Sub\,(\Gamma\ext A)\,(\Gamma\ext A)$, we recover
  the \emph{weakening substitution} $\p : \Sub\,(\Gamma\ext A)\,\Gamma$ and the \emph{zero variable}
  $\q : \Tm_{i,j}\,(\Gamma\ext A)\,(A[\p])$.
\item
  By weakening $\q$, we recover a notion of variables as De Bruijn
  indices. In general, the $n$-th De Bruijn index is defined as $\q[\p^{n}]$,
  where $\p^{n}$ denotes $n$-fold composition.
\item
  By moving right-to-left along $\Sub\,\Gamma\,(\Delta\ext A) \simeq ((\sigma :
  \Sub\,\Gamma\,\Delta) \times \Tm_{i,j}\,\Gamma\,(A[\sigma]))$, we recover the
  operation which extends a morphism with a term. In the syntax (initial model),
  this justifies the view of $\Sub$ as a type of lists of terms, i.e.\ a
  parallel substitutions. We denote the extension operation as $(\sigma,\,t)$
  for $\sigma : \Sub\,\Gamma\,\Delta$.
\end{itemize}

\begin{notation}
De Bruijn indices are rather hard to read, so we will often use nameful notation
for binders and substitutions. For example, we may write $\Gamma \ext (x : A)
\ext (y : B)$ for a context, and subsequently write $B[y \mapsto t]$ for
substituting the $x$ variable for some term $t : \Tm_{i,j}\,\Gamma\,A$. Using
nameless notation, we would instead have $B : \Ty_{i,j}\,(\Gamma \ext A)$ and
$B[\id,\,t]$; here we recover single substitution by extending the identity
substitution $\id : \Sub\,\Gamma\,\Gamma$ with $t$.

We may also use implicit weakening: if a type or term is in a $\Gamma$ context,
we may use it in an extended $\Gamma \ext A$ context without marking the
weakening substitution.
\end{notation}

\begin{definition} A \textbf{model of 2LTT} is a model of basic 2LTT
which supports certain type formers. For the sake of brevity, we only present
our results for a small collection of type formers. However, we will argue that
our results easily extend to any general notion of inductive type formers. We
specify type formers in the following.  We assume that all type formers are
natural, i.e.\ stable under substitution.
\begin{itemize}
\item \emph{Universes}. For each $i$ and $j$, we have a Coquand-style universe
  \cite{TODO} in $\Ty_{i,j}$. This consists of $\U_{i,j} : \Ty_{i,j+1}\,\Gamma$,
  together with $\El : \Tm_{i,j+1}\,\Gamma\,\U_{i,j} \to \Ty_{i,j}\,\Gamma$ and $\Code$, where $\Code$
  and $\El$ are inverses.
\item \emph{$\Sigma$-types}. We have $\Sigma\,(x : A)\,B : \Ty_{i,j}\,\Gamma$ for $A : \Ty_{i,j}\,\Gamma$
  and $B : \Ty_{i,j}\,(\Gamma \ext (x : A))$, together with a natural isomorphism of pairing and projections:
  \[ \Tm_{i,j}\,\Gamma\,(\Sigma\,A\,B) \simeq ((t : \Tm_{i,j}\,\Gamma\,A) \times \Tm_{i,j}\,\Gamma\,(B[x\mapsto t]) \]
  We write $(t,\,u)$ for pairing and $\fst$ and $\snd$ for projections.
\item \emph{Function types}. We have $\Pi\,(x : A)\,B : \Ty_{i,j}\,\Gamma$ for $A :
  \Ty_{i,j}\,\Gamma$ and $B : \Ty_{i,j}\,(\Gamma \ext (x : A))$, together with
  $\app : \Tm_{i,j}\,\Gamma\,(\Pi\,A\,B) \to \Tm_{i,j}\,(\Gamma \ext (x :
  A))\,B$ and its inverse $\lam$.
\item \emph{Natural numbers}. We have $\Nat_{i,j} : \Ty_{i,j}\,\Gamma$, $\zero_{i,j} : \Tm_{i,j}\,\Gamma\,\Nat_{i,j}$,
  and $\suc_{i,j} : \Tm_{i,j}\,\Gamma\,\Nat_{i,j} \to \Tm_{i,j}\,\Gamma\,\Nat_{i,j}$. The eliminator is the following.
  \begin{alignat*}{5}
    & \NatElim :\hspace{0.5em}
                &&  (P &&: \Ty_{i,k}\,(\Gamma \ext (n : \Nat_{i,j}) \\
    &           &&  (z &&: \Tm_{i,k}\,\Gamma\,(P[n\mapsto \zero_{i,j}])) \\
    &           &&  (s &&: \Tm_{i,k}\,(\Gamma \ext (n : \Nat_{i,j}) \ext (\mit{pn} : P[n\mapsto n]))\,(P[n\mapsto \suc_{i,j}\,n]))\\
    &           &&  (t &&: \Tm_{i,j}\,\Gamma\,\Nat_{i,j}) \\
    &\hspace{3em}\to        &&&& \Tm_{i,j}\,\Gamma\,(P[n\mapsto t]))
  \end{alignat*}
  We also have the $\beta$-rules:
  \begin{alignat*}{5}
    & \NatElim\,P\,z\,s\,\zero_{i,j}     &&= z \\
    & \NatElim\,P\,z\,s\,(\suc_{i,j}\,t) &&= s[n \mapsto t,\,\mit{pn} \mapsto \NatElim\,P\,z\,s\,t]
  \end{alignat*}
  Note that we can eliminate from a level $j$ to any level $k$.
\end{itemize}
\end{definition}

\begin{definition}
The \textbf{syntax of 2LTT} is defined to be the initial model of 2LTT. The syntax
can be assumed to exist, since the notion of model is algebraic; the
specification can be expressed using a suitable notion of algebraic signatures,
such as one for essentially algebraic theories \cite{TODO} or as a finitary
quotient inductive-inductive signature \cite{TODO}. The initiality of syntax
directly yields a notion of \emph{recursion on syntax}, and it can be shown that
initiality also implies an \emph{induction principle} \cite{TODO}.
\end{definition}

The examples in Section \ref{TODO} can be faithfully represented in the formal
syntax. The only notable difference is that the formal syntax uses Coquand-style
universes instead of Russell-style ones. We use the former because it is much
easier to model in the semantics in Section \ref{TODO}. We note though that it
is straightforward to elaborate from a Russell-style surface syntax to the
formal version, by inserting $\El$ and $\Code$ as required.

\subsubsection{Comparison to Annekov et al.}
Comparing our models to the primary reference on 2LTT \cite{twolevel}, the main
difference is the handling of ``sizing'' levels. In ibid.\ there is a cumulative
lifting from $\Ty_{i,j}$ to $\Ty_{i,j+1}$, which we do not assume. Instead,
we allow elimination from $\Nat_{i,j}$ into any $k$ level. This means that we
can manually define lifting maps from $\Nat_{i,j}$ to $\Nat_{i,j+1}$ by
elimination. This is more similar to e.g.\ Agda, where we do not have cumulativity,
but we can define explicit lifting from $\Nat_j : \Set_j$ to $\Nat_k : \Set_k$.

In \cite{twolevel}, ``two-level type theory'' specifically refers to the setup
where the object-level is a homotopy type theory and the meta level is an
extensional type theory. In contrast, we allow a wider range of setups under the
2LTT umbrella. Annekov et al.\ also considers a range of additional
strengthenings and extension of 2LTT \cite[Section~2.4]{twolevel}, most of which
are of use in synthetic homotopy theory. We do not assume any of these, and
stick to the most basic formulation of 2LTT.

\subsection{Models and Syntax of the Object Theory}

We also need to specify the object theory, which serves as the output language
of staging. In general, the object theory corresponding to a particular flavor
of 2LTT is simply the type theory that supports only the object-level $\Ty_{0,j}$
hierarchy with its type formers.

\begin{definition}
A \textbf{model of the object theory} is a category-with-families, with types and
terms indexed over $j \in \mbb{N}$, supporting Coquand-style universes $\U_j$,
type formers $\Pi$, $\Sigma$ and $\Nat_j$, with elimination from $\Nat_j$ to
any level $k$.
\end{definition}

\begin{definition}
Like before, the \textbf{syntax of the object theory} is the initial model.
\end{definition}

\begin{notation}
From now on, by default we use $\Con$, $\Sub$, $\Ty$ and $\Tm$ to refer to sets
in the syntax of 2LTT. We use $\mbbo$ to refer to the object syntax, and
$\Con\ob$, $\Sub\ob$, $\Ty\ob$ and $\Tm\ob$ to refer to its underlying sets.
\end{notation}

\begin{definition}
We have an \textbf{embedding} of object syntax syntax into 2LTT syntax, consisting
of the following functions:
\begin{alignat*}{4}
  & \emb{\blank} : \Con\ob \to \Con\\
  & \emb{\blank} : \Sub\ob\,\Gamma\,\Delta \to \Sub\,\emb{\Gamma}\,\emb{\Delta}\\
  & \emb{\blank} : \Ty_{\mbbo j}\,\Gamma \to \Ty_{0,j}\,\emb{\Gamma}\\
  & \emb{\blank} : \Tm_{\mbbo j}\,\Gamma\,A \to \Tm_{0,j}\,\emb{\Gamma}\,\emb{A}
\end{alignat*}
Embedding maps all type and term formers to the corresponding ones in 2LTT, and
strictly preserves all structure.
\end{definition}

\section{The Staging Algorithm}

In this section we specify what we mean by staging, then define a staging
algorithm.

\begin{definition}
A \textbf{staging algorithm} consists of two functions:
  \begin{alignat*}{3}
    & \Stage : \Ty_{0,j}\,\emb{\Gamma} \to \Ty_{\mbbo j}\,\Gamma \\
    & \Stage : \Tm_{0,j}\,\emb{\Gamma}\,A \to \Tm_{\mbbo j}\,\Gamma\,(\Stage\,A)
  \end{alignat*}
Note that we can stage open types and terms as long as their contexts are purely
object-level. By \emph{closed staging} we mean staging only for closed types and
terms.
\end{definition}
\begin{definition}
  A staging algorithm $\Stage$ is \textbf{correct} if the following properties hold:
  \begin{itemize}
  \item \emph{Soundness:} $\emb{\Stage\,A} = A$ and $\emb{\Stage\,t} = t$.
  \item \emph{Completeness:} staging respects definitional equality.
  \item \emph{Stability:} $\Stage\,\emb{A} = A$ and $\Stage\,\emb{t} = t$.
  \end{itemize}
\end{definition}
We make some remarks on correctness. First, we get completeness for free in our
algebraic setup, since \emph{all} functions definable in the metatheory must
respect definitional equality.

Second, note that soundness and stability together is the statement that
embedding is a bijection on types and terms. This is a statement of
\emph{conservativity} of 2LTT over the object theory. In \cite{TODO} a
significantly weaker conservativity theorem is shown, which expresses that there
exists a function from $\Tm_{0,j}\,\emb{\Gamma}\,\emb{A}$ to $\Tm_{\mbbo
  j}\,\Gamma\,A$.

Lastly, we note that the correctness terminology is borrowed from previous works
on normalization-by-evaluation, where the soundness-completeness-stability trio
has been used several times \cite{TODO}. In \cite{cubicalnbe}, stability is
instead called ``idempotence''.

\subsection{The Presheaf Model}

We observe that the presheaf model described in \cite[Section~2.5.3]{twolevel}
immediately yields a closed staging algorithm, by evaluation of 2LTT types and
terms in the model. In this model, contexts are presheaves over the syntactic
category of the object theory. We call the model $\hato$ and denote its
components by putting hats on 2LTT components. We summarize the key components
in the following.

\begin{notation}
In this section, we switch to naming elements of $\Cono$ as $a$, $b$ and $c$,
and elements of $\Subo$ as $f$, $g$, and $h$, to avoid name clashing with
contexts and substitutions in the presheaf model.
\end{notation}

\subsubsection{The syntactic category and the meta-level fragment}

\begin{definition} $\bs{\wh{\Con} : \Set_{\omega+1}}$ is defined as the set of presheaves
over $\mbbo$. $\Gamma : \wh{\Con}$ has an action on objects $|\Gamma| :
\Cono \to \Set_\omega$ and an action on morphisms $\blank[\blank] : |\Gamma|\,b
\to \Subo\,a\,b \to |\Gamma|\,a$, such that $\gamma[\id] = \gamma$ and
$\gamma[f\circ g] = \gamma[f][g]$.
\begin{notation}
We reuse the substitution notation $\blank[\blank]$ for the action on morphisms.
Also, we use lowercase $\gamma$ and $\delta$ to denote elements of $|\Gamma|\,a$
and $|\Delta|\,a$ respectively.
\end{notation}
\end{definition}
\begin{definition} $\bs{\hSub\,\Gamma\,\Delta : \Set_\omega}$ is the set of natural transformations
from $\Gamma$ to $\Delta$. $\sigma : \hSub\,\Gamma\,\Delta$ has action
$|\sigma| : \{a : \Cono\} \to |\Gamma|\,a \to |\Delta|\,a$ such that
$|\sigma|\,(\gamma[f]) = (|\sigma|\,\gamma)[f]$.
\end{definition}

\begin{definition}
$\bs{\hTy_{1,j}\,\Gamma : \Set_\omega}$ is the set of \emph{displayed presheaves}
over $\Gamma$; see e.g.\ \cite{TODO}. This is equivalent to the set of
presheaves over the category of elements of $\Gamma$, but it is usually more
convenient in calculations. An $A : \hTy\,\Gamma$ has an action on objects
$|A| : \{a : \Cono\} \to |\Gamma|\,a \to \Set_j$ and an action on
morphisms $\blank[\blank] : |A|\,\gamma \to (f : \Subo\,a\,b) \to
|A|\,(\gamma[f])$, such that $\alpha[\id] = \alpha$ and $\alpha[f \circ g] = \alpha[f][g]$.

\begin{notation}
  We write $\alpha$ and $\beta$ respectively for elements of $|A|\,\gamma$ and
  $|B|\,\gamma$.
\end{notation}
\end{definition}

\begin{definition}
  $\bs{\hTm_{1,j}\,\Gamma\,A : \Set_\omega}$ is the set of sections of the
  displayed presheaf $A$. This can be viewed as a dependently typed analogue of
  a natural transformation.  A $t : \hTm_{1,j}\,\Gamma\,A$ has action $|t| :
  \{a\} \to (\gamma : |\Gamma|\,a) \to |A|\,\gamma$ such that $|t|\,(\gamma[f])
  = (|t|\,\gamma)[f]$.
\end{definition}

%% \begin{definition}
%% $\bs{\wh{\U}_{1,j} : \hTy_{1,j+1}\,\Gamma}$ is defined by
%% $|\wh{\U}_{1,j}|\,\{a\}\,\_ = \hTy_{1,j}\,(\y\,a)$, and defining action on
%% morphisms by $\hTy$ substitution. Note that the universe levels check out in
%% this definition, since by unfolding $\hTy_{1,j}\,(\y\,a)$ we can show that it
%% is in $\Set_j$.
%% \end{definition}

We also look at he empty context and context extension with meta-level types,
as these will appear in subsequent definitions.

\begin{definition}
$\bs{{\widehat{\emptycon}} : \hCon}$ is defined as the presheaf
which is constantly $\top$, i.e.\ $|\wh{\emptycon}|\,\_ = \top$.
\end{definition}

\begin{definition} For $A : \hTy_{1,j}\,\Gamma$, we define $\bs{\Gamma\,\,\wh{\ext}\,A}$
pointwise by $|\Gamma\,\,\wh{\ext}\,A|\,a := (\gamma : |\Gamma|\,a) \times |A|\,\gamma$
and $(\gamma,\,\alpha)[f] := (\gamma[f],\,\alpha[f])$.
\end{definition}

Using the above definitions, we can model the syntactic category of 2LTT, and
also the meta-level family structure and all meta-level type formers. For an
exposition in previous literature, see \cite{TODO} and \cite{TODO}.

\subsubsection{The object-level fragment}

We move on to modeling the object-level syntactic fragment of 2LTT. We make some
preliminary definitions. First, note the types in the object theory yield a
presheaf, and terms yield a displayed presheaf over them; this immediately
follows from the specification of a family structure in a cwf. Hence, we do a
bit of a name overloading, and have $\Ty_{\mbbo j} : \hCon$ and $\Tm_{\mbbo j} : \hTy\,\Ty_{\mbbo j}$.

\begin{definition}
$\bs{\hTy_{0,j}\,\Gamma : \Set_\omega}$ is defined as $\hSub\,\Gamma\,\Ty_{\mbbo j}$,
and $\bs{\hTm_{0,j}\,\Gamma\,A : \Set_\omega}$ is defined as $\hTm\,\Gamma\,(\Tm_{\mbbo j}[A])$.
\end{definition}

For illustration, if $A : \hTy_{0,j}\,\Gamma$, then $A :
\hSub\,\Gamma\,\Ty_{\mbbo j}$, so $|A| : \{a : \Cono\} \to |\Gamma|\,a \to
\Ty_{\mbbo j}\,a$. In other words, the action of $A$ on objects maps a semantic
context to a syntactic object-level type. Likewise, for $t :
\hTm_{0,j}\,\Gamma\,A$, we have $|t| : (\gamma : |\Gamma|\,a) \to \Tm_{\mbbo
  j}\,a\,(|A|\,\gamma)$, so we get a syntactic object-level term as output.

\begin{definition} For $A : \Ty_{0,j}\,\Gamma$, we define $\bs{\Gamma\,\,\wh{\ext}\,A}$
as $|\Gamma\,\,\wh{\ext}\,A|\,a := (\gamma : |\Gamma|\,a) \times \Tm_{\mbbo
  j}\,a\,(|A|\,\gamma)$ and $(\gamma,\,t)[f] := (\gamma[f],\,t[f])$. Thus,
extending a semantic context with an object-level type appends an object-level
term.
\end{definition}

Using the above definitions and following \cite{TODO}, we can model all type
formers in $\hTy_{0,j}$. Intuitively, that is because $\hTy_{0,j}$ and
$\hTm_{0,j}$ return types and terms, so we can reuse the type and term formers
in the object theory.

\subsubsection{Lifting}

\begin{definition}
$\bs{\wh{\Lift} A : \hTy_{1,j}\,\Gamma}$ is defined as $\Tm_{\mbbo j}[A]$. With
  this, we get that $\hTm_{0,j}\,\Gamma\,A$ is equal to
  $\hTm_{1,j}\,\Gamma\,(\wh{\Lift}\,A)$. Hence, we can define both quoting and
  splicing as identity functions in the model.
\end{definition}

\subsection{Closed Staging}

\begin{definition} The \textbf{evaluation morphism, denoted $\bs{\ev}$} is
a 2LTT model morphism from the syntax of 2LTT to $\hato$, arising from the initiality
of the syntax. In other words, $\ev$ is defined by recursion on the syntax and
strictly preserves all structure.
\end{definition}

Note that $\wh{\emptycon}$ is defined as the terminal presheaf, which is
constantly $\top$. This leads us to the following definition.

\begin{definition} \textbf{Closed staging} is defined as follows.
\begin{alignat*}{4}
  & \Stage : \Ty_{0,j}\,\emptycon \to \Ty_{\mbbo j}\,\emptycon \hspace{2em} && \Stage : \Tm_{0,j}\,\emptycon\,A \to \Tm_{\mbbo j}\,\emptycon\,(\Stage\,A) \\
  & \Stage\,A := |\ev\,A|\,\{\emptycon\}\,\tt && \Stage\,t := |\ev\,t|\,\{\emptycon\}\,\tt
\end{alignat*}
This is well-typed, since $|\ev\,A|\{\emptycon\} : |\ev\,\emptycon|\,\emptycon \to \Ty_{\mbbo
  j}\,\emptycon$, so $|\ev\,A|\{\emptycon\} : \top \to \Ty_{\mbbo
  j}\,\emptycon$, and $|\ev\,t|\,\{\emptycon\} : \top \to \Tm_{\mbbo
  j}\,\emptycon\,(|\ev\,A|\,\tt)$.
\end{definition}

What about general (open) staging though? Given $A : \Ty_{0,j}\,\emb{\Gamma}$,
we get $|\ev\,A|\,\{\Gamma\} : |\emb{\Gamma}|\,\Gamma \to \Ty_{\mbbo j}$. We
need an element of $|\emb{\Gamma}|\,\Gamma$, in order to extract an object-level
type. Such ``generic'' semantic contexts should be possible to construct, because
elements of $|\emb{\Gamma}|\,\Gamma$ are essentially lists of object-level terms
in $\Gamma$, by Definition \ref{TODO}, so $|\emb{\Gamma}|\,\Gamma$ should be
isomorphic to $\Subo\,\Gamma\,\Gamma$.  It turns out that this falls out from
the stability proof of $\ev$.

\subsection{Stability and Open Staging}

Stability means that staging an object-level construction does nothing. We
define a family of functions $\blank^P$ by induction on object syntax. The
induction motives are as follows.
\begin{alignat*}{4}
  & (\Gamma : \Cono)^P                    &&: |\ev\,\emb{\Gamma}|\,\Gamma\\
  & (\sigma : \Subo\,\Gamma\,\Delta)^P    &&: \Delta^P[\sigma] = |\ev\,\emb{\sigma}|\,\Gamma^P\\
  & (A      : \Ty_{\mbbo j}\,\Gamma)^P      &&: A = |\ev\,\emb{A}|\,\Gamma^P\\
  & (t      : \Tm_{\mbbo j}\,\Gamma\,A)^P   &&: t = |\ev\,\emb{t}|\,\Gamma^P
\end{alignat*}
We look at the interpretation of contexts.
\begin{itemize}
\item  For $\emptycon^P$, we need an element of $|\ev\,\emb{\emptycon}|\,\emptycon$, hence
  an element of $\top$, so we define $\emptycon^P$ as $\tt$.
\item
  For $(\Gamma \ext A)^P$, we need an element of
    \[(\gamma : |\ev{\emb{\Gamma}}|\,(\Gamma\ext A)) \times \Tm_{\mbbo j}\,(\Gamma \ext A)\,(|\ev\emb{A}|\,\gamma).\]
  We have $\Gamma^P : |\ev{\emb{\Gamma}}|\,\Gamma$, which we can weaken as
  $\Gamma^P[\p] : |\ev{\emb{\Gamma}}|\,(\Gamma \ext A)$, so we set the first
  projection of the result as $\Gamma^P[\p]$. For the second projection, the
  goal type can be simplified as follows:
  \begin{alignat*}{4}
    &       &&\Tm_{\mbbo j}\,(\Gamma \ext A)\,(|\ev\emb{A}|\,(\Gamma^P[\p])) && \\
    & =\,\, &&\Tm_{\mbbo j}\,(\Gamma \ext A)\,((|\ev\emb{A}|\,\Gamma^P)[\p]) &&\hspace{2em} \text{by naturality of $\ev\emb{A}$}\\
    & =\,\, &&\Tm_{\mbbo j}\,(\Gamma \ext A)\,(A[\p])                        &&\hspace{2em} \text{by $A^P$}
  \end{alignat*}
  We have the zero de Bruijn variable $\q : \Tm_{\mbbo j}\,(\Gamma \ext
  A)\,(A[\p])$.  Hence, we define $(\Gamma \ext A)^P$ as $(\Gamma^P[\p],\,\q)$.
\end{itemize}

Thus, a generic semantic context $\Gamma^P : |\ev\emb{\Gamma}|\,\Gamma$ is just
a list of variables, corresponding to the identity substitution $\id :
\Subo\,\Gamma\,\Gamma$ which maps each variable to itself.

The rest of the $\blank^P$ interpretation is straightforward and we omit it
here. In particular, we do not have to show preservation of any definitional
equality, since types, terms and substitutions are all interpreted as
proof-irrelevant equations.

\begin{definition} We define \textbf{open staging} as follows.
\begin{alignat*}{4}
  & \Stage : \Ty_{0,j}\,\emb{\Gamma} \to \Ty_{\mbbo j}\,\Gamma \hspace{2em} && \Stage : \Tm_{0,j}\,\emb{\Gamma}\,A \to \Tm_{\mbbo j}\,\Gamma\,(\Stage\,A) \\
  & \Stage\,A := |\ev\,A|\,\Gamma^P && \Stage\,t := |\ev\,t|\,\Gamma^P
\end{alignat*}
\end{definition}

\begin{theorem} Open staging is stable.
\end{theorem}
\begin{proof} For $A : \Ty_{\mbbo j}$, $\Stage\,\emb{A}$ is by definition $|\ev\,\emb{A}|\,\Gamma^P$,
  hence by $A^P$ it is equal to $A$. Likewise, $\Stage\,\emb{t}$ is equal to $t$ by $t^P$.
\end{proof}

\subsection{Implementation and Efficiency}

We have defined staging in a constructive metatheory, so it is possible to
extract an implementation from it. We discuss its operational behavior and
potential optimizations in the following. First, consider the computational
part of evaluation:
\begin{alignat*}{3}
  & |\blank| \circ \ev : \Ty_{1 j}\,\Gamma \to \{\Delta : \Cono\} \to |\ev\,\Gamma|\,\Delta \to \Set_j\\
  & |\blank| \circ \ev : \Ty_{0 j}\,\Gamma \to \{\Delta : \Cono\} \to |\ev\,\Gamma|\,\Delta \to \Ty_{\mbbo j}\,\Delta\\
  & |\blank| \circ \ev : \Tm_{1 j}\,\Gamma\,A \to \{\Delta : \Cono\} \to (\gamma : |\ev\,\Gamma|\,\Delta) \to
    |\ev\,A|\,\gamma \\
  & |\blank| \circ \ev : \Tm_{0 j}\,\Gamma\,A \to \{\Delta : \Cono\} \to (\gamma : |\ev\,\Gamma|\,\Delta) \to
    \Tm_{\mbbo\,j}\,\Delta\,(|\ev\,A|\,\gamma)
\end{alignat*}
Operationally, we interpret syntactic types or terms in a semantic environment
$\gamma : |\ev\,\Gamma|\,\Delta$. These environments are lists containing
object-level terms and semantic values. The meta-level semantic values are
represented using meta-level inductive types and the function space of the
metatheory.
\begin{itemize}
  \item $|\Nat_1|\,\_$ is simply the set of meta-level natural numbers.
  \item $|\Sigma_1\,A\,B|\,\gamma$ is simply a set of pairs of values.
  \item In the non-dependent case, a function type $A\,\wh{\to}\,B$ is defined as the presheaf
        exponential \cite{TODO}. The computational part of $|A\,\wh{\to}\,B|\,\{\Delta\}\,\gamma$
        is given by a function with type
        \[ (\Theta : \Cono) \to (\sigma : \Subo\,\Theta\,\Delta) \to |A|\,(\gamma[\sigma]) \to |B|\,(\gamma[\sigma])        \]
        This may be also familiar as the semantic implication from the Kripke
        semantics of intuitionistic logics. Whenever we evaluate a function
        application, we supply an extra $\id : \Subo\,\Delta\,\Delta$. This may
        incur cost via the $\gamma[\id]$ restrictions in a naive
        implementation, but this is easy to optimize, by introducing a formal
        representation of $\id$, such that $\gamma[\id]$ immediately computes to
        $\gamma$. The dependent case is analogous operationally.
\end{itemize}
In summary, in the meta-level fragment of 2LTT, $\ev$ yields a reasonably efficient
computation of closed values, which reuses functions from the ambient
metatheory. Alternatively, instead of using ambient functions, we could use our
own implementation of function closures during staging.

In contrast, we have a bit of an efficiency problem in the handling of object-level
terms and binders: whenever we go under an object-level binder we have to weaken
the current semantic context. In other words, when moving from $\Delta$ to
$\Delta \ext A$, we have to shift $\gamma : |\ev\,\Gamma|\,\Delta$ to
$\gamma[\p] : |\ev\,\Gamma|\,(\Delta \ext A)$. This cannot be avoided by an easy
optimization trick.

The exact same efficiency issue occurs in formal presheaf-based
normalization-by-evaluation \cite{TODO}. However, in practical implementations
\cite{TODO} this issue can be fully solved by using De Bruijn \emph{levels} in
the semantic domain, thus arranging that weakening has no operational cost on
semantic values. This corresponds to Abel's ``liftable terms'' presentation
\cite{TODO}. We can use the same solution in staging, by using de Bruijn levels
and delayed weakenings in object-level terms during staging. It is a somewhat
unfortunate mismatch that indices are far more convenient in formalization, but
levels are more efficient in practice.

In our prototype implementation, we use the above optimization with De Bruijn
levels, and we also switch to \emph{syntax-directed} staging. In the formal
staging algorithm, since types are tracked, quotes and splices have no
operational significance; recall that they are interpreted as identity functions
in Definition \ref{TODO}. However, it seems more efficient to not compute types
at all and instead rely on quotes and splices to keep track of the current
stage.

\subsubsection{Caching} In a production-strength implementation we would need some
caching mechanism, to avoid excessive duplication of function code. For example,
if we use the staged $\mit{map}$ function multiple times with the same type and
function arguments, we do not want to generate separate code for each
usage. De-duplicating object-level code with \emph{function types} is
usually safe, since function bodies become closed top-level code after closure
conversion. We leave this to future work.

\section{Yoneda, Representability and Generativity}

\section{Soundness of Staging}

In this section we prove soundness of staging. We build a proof-relevant logical
relation between the evaluation morphism $\ev$ and a \emph{restriction}
morphism, which restricts 2LTT syntax to object-level contexts. The relational
interpretations of $\Ty_{0,j}$ and $\Tm_{0,j}$ will yield the soundness
property.

\subsection{Working in $\hat{\mbbo}$}

We have seen that 2LTT can be modeled in the presheaf topos
$\hato$. Additionally, $\hato$ supports all type formers of extensional type
theory \cite{TODO}, and certain other structures which are stable under
object-level substitution. Since almost all constructions in this section must
be stable under object-level substitution, it makes sense to work internally to
$\hato$. This technique has been previously used in normalization proofs
\cite{TODO} and also in the metatheory of cubical type theories \cite{TODO}.

When we work internally in a model of a type theory, we do not explicitly refer
to contexts, types, and substitutions. For example, when working in Agda, we do
not refer to Agda's typing contexts. Instead, we only work with terms, and use
functions and universes to abstract over types and semantic contexts. Hence, we
have to convert along certain isomorphisms when we switch between the internal
and external views. In the following, we summarize features in $\hato$ and also
the internal-external conversions.

We write $\whset_j$ for ordinal-indexed Russell-style universes. Formally, we
have Coquand-style universes, but for the sake of brevity we omit $\El$ and
$\Code$ from internal syntax. Universes are cumulative, and closed under $\Pi$,
$\Sigma$, extensional identity $\blank\!=\!\blank$ and inductive types. We use
the same conventions as in Notation \ref{TODO}.

\noindent The basic scheme for internalization is as follows:
\begin{alignat*}{3}
  & \Gamma : \hCon                  && \hspace{1em}\text{is internalized as}\hspace{1em} \Gamma : \whset_\omega\\
  & \sigma : \hSub\,\Gamma\,\Delta  && \hspace{1em}\text{is internalized as}\hspace{1em} \sigma : \Gamma \to \Delta\\
  & A : \hTy_{1,j}\,\Gamma           && \hspace{1em}\text{is internalized as}\hspace{1em} A      : \Gamma \to \whset_j \\
  & t : \hTm_{1,j}\,\Gamma\,A        && \hspace{1em}\text{is internalized as}\hspace{1em} t      : (\gamma : \Gamma) \to A\,\gamma
\end{alignat*}
We internalize object-level types and terms as $\Ty_{\mbbo j} : \whset_0$ and
$\Tm_{\mbbo j} : \Ty_{\mbbo j} \to \whset_0$.  We also use an internal view of
$\ev$ which maps 2LTT syntax to internal values; i.e.\ we compose $\ev$ with
internalization. Below, note that the input to $\ev$ is \emph{external}, so we
mark $\ev$ as being parametrized by external input.
\begin{alignat*}{3}
  &\ev\,(\Gamma &&: \Con)                      &&: \whset_\omega \\
  &\ev\,(\sigma &&: \Sub\,\Gamma\,\Delta)      &&: \ev\,\Gamma \to \ev\,\Delta \\
  &\ev\,(A      &&: \Ty_{1,j}\,\Gamma)          &&: \ev\,\Gamma \to \whset_j\\
  &\ev\,(t      &&: \Tm_{1,j}\,\Gamma)          &&: (\gamma : \ev\,\Gamma) \to \ev\,A\,\gamma\\
  &\ev\,(A      &&: \Ty_{0,j}\,\Gamma)          &&: \ev\,\Gamma \to \Ty_{\mbbo_j}\\
  &\ev\,(t      &&: \Tm_{0,j}\,\Gamma\,A)       &&: (\gamma : \ev\,\Gamma) \to \Tm_{\mbbo j}\,(\ev\,A\,\gamma)
\end{alignat*}
The purely object-level syntactic fragment of 2LTT can be internalized as
follows. We define externally the presheaf of object-level types as
$|\Ty_{0,j}|\,a := \Ty_{0,j}\,\emb{a}$, and the displayed presheaf of
object-level terms over $\Ty_{0,j}$ as $|\Tm_{0,j}|\,\{a\}\,A :=
  \Tm_{0,j}\,\emb{a}\,A$. Hence, internally we have $\Ty_{0,j} : \whset_0$ and
  $\Tm_{0,j} : \Ty_{0,j} \to \whset_0$.

Now, the embedding operation $\emb{\blank}$ can be also internalized on types
and terms, as $\emb{\blank} : \Ty_{\mbbo j} \to \Ty_{0,j}$, and $\emb{\blank}
:\Tm_{\mbbo\,j}\,A \to \Tm_{0,j}\,\emb{A}$.

\subsection{The Restriction Morphism}

We define a family of functions $\re$ from the 2LTT syntax to objects in
$\hato$. This will be related to the evaluation morphism $\ev$ subsequently, in
the relational interpretation. In short, $\re$ restricts 2LTT syntax so that it
can only depend on object-level contexts, i.e.\ contexts given as
$\emb{\Gamma}$. Below, we specify the types of the $\re$ operations internally,
and the $|\blank|$ components of the operations externally. The naturality of
$|\blank|$ is straightforward in each case.
\begin{alignat*}{5}
  & \re\,(\Gamma : \Con) : \whset_0 && \re\,(\sigma : \Sub\,\Gamma\,\Delta) : \re\,\Gamma \to \re\,\Delta \\
  & |\re\,\Gamma|\,a := \Sub\,\emb{a}\,\Gamma && |\re\,\sigma|\,\gamma := \sigma \circ \gamma \\
  & \\
  & \re\,(A : \Ty_{1,j}\,\Gamma) : \re\,\gamma \to \whset_0 && \re\,(t : \Tm_{1,j}\,\Gamma\,A) : (\gamma : \re\,\Gamma) \to \re\,A\,\gamma \\
  & |\re\,A|\,\{a\}\,\gamma := \Tm_{1,j}\,\emb{a}\,(A[\gamma])\hspace{2em} && |\re\,t|\,\gamma := t[\gamma]\\
  & \\
  & \ev\,(A : \Ty_{0,j}\,\Gamma) : \re\,\Gamma \to \Ty_{0,j}\hspace{2em} && \ev\,(t : \Tm_{0,j}\,\Gamma\,A) : (\gamma : \re\,\Gamma) \to \Tm_{0,j}\,(\re\,A\,\gamma)\\
  & |\ev\,A|\,\gamma := A[\gamma] && |\ev\,t|\,\gamma := t[\gamma]
\end{alignat*}

\subsubsection{Preservation properties of $\re$}
First, we note that $\re$ strictly preserves $\id$, $\blank\circ\blank$ and
type/term substitution, and it preserves $\emptycon$ and $\blank\ext_1\blank$ up
to isomorphism. We have the following isomorphisms internally to $\hato$:
\begin{alignat*}{3}
  & \re_{\emptycon} &&: \re\,\emptycon \simeq \top \\
  & \re_{\ext_1}   &&: \re\,(\Gamma \ext_1 A) \simeq ((\gamma : \re\,\Gamma) \times \re\,A\,\gamma)
\end{alignat*}
\begin{notation}
  When we have an isomorphism $f : A \simeq B$, we may write $f$ for the
  function in $A \to B$, and $f^{-1} : B \to A$ for its inverse.
\end{notation}
The above preservation properties mean that $\re$ is a \emph{pseudomorphism} in
the sense of \cite{gluing}, between the syntactic cwf given by
$(\Ty_{1,j},\,\Tm_{1,j})$ and the corresponding cwf structure in $\hato$. In
ibid.\ there is an analysis of such cwf morphisms, from which we obtain the
following additional preservation properties:
\begin{itemize}
\item Meta-level negative type formers are preserved up to isomorphism. In particular, we have
  \[ \re_{\Sigma} : \re\,(\Sigma\,A\,B)\,\gamma
  \simeq ((\alpha : \re\,A) \times \re\,B\,(\rexti\,(\gamma,\,\alpha))). \]
  To provide a higher-level explanation, $\re\,(\Sigma\,A\,B)\,\gamma$ is
  essentially the presheaf of 2LTT terms with type $\Sigma_1\,A\,B$, but the
  terms are restricted to object-level contexts. However, we can still perform
  pairing and projection on such restricted terms; hence the preservation property.

\item Meta-level $\Pi$ types and universes are preserved in a lax way. For $\Pi$, we have
  \[ \re_{\app} : \re\,(\Pi\,A\,B)\,\gamma \to (\alpha : \re\,A\,\gamma) \to \re\,B\,(\rexti\,(\gamma,\,\alpha))
  \]
  such that $\re_{\app}\,(\re\,(\lam\,t)\,\gamma)\,\alpha =
  \re\,t\,(\rexti\,(\gamma,\,\alpha))$. In this case, we can apply a restricted
  term with a $\Pi$-type to a restricted term, but we cannot do
  lambda-abstraction, because that would require extending the context with a
  meta-level binder.  For $\U_{1,j}$, we have
  \[
    \re_{\Code} : \re\,\U_{1,j}\,\gamma \to \whset_j
  \]
  such that $\re_{\Code}\,(\re\,(\El\,t)\,\gamma) = \re\,t\,\gamma$. We skip
  $\Code$ on the right hand side here, since we assume Russell-style notation
  internally to $\hato$, i.e.\ we omit $\Code$ and $\El$. Here, we only
  have lax preservation simply because $\whset_j$ is much larger than
  the the set of syntactic 2LTT types, so not every semantic $\whset_j$
  have a syntactic representation.

\item Meta-level positive (inductive) types are preserved in an oplax way. In the case of
  natural numbers, we have
  \[ \re_{\mbb{N}} : \mbb{N} \to \re\,\Nat_1\,\gamma. \]
  This is a ``serialization'' map: from a metatheoretic natural number we
  compute a numeral as a closed 2LTT term. This works analogously for any
  inductive type. The restriction to object-level contexts is irrelevant when
  computing closed terms.
\end{itemize}
\paragraph{Action on lifting} We have an isomorphism $\re\,(\Lift\,A)\,\gamma \simeq
\Tm_{0,j}\,(\re\,A\,\gamma)$. This is given by quoting and splicing: we convert
between restricted meta-level terms with type $\Lift\,A$ and restricted
object-level terms with type $A$. Hence, we simply write the components of this
isomorphism as $\spl\blank$ and $\qut{\blank}$, as internal analogues of the
external operations. With this, we also have $\re\,\qut{t}\,\gamma =
\qut{\re\,t\,\gamma}$ and $\re\,\spl\,t\,\gamma = \spl(\re\,t\,\gamma)$.

\paragraph{Action on $\blank\ext_0\blank$} Lastly, $\blank\ext_0\blank$ is preserved
up to isomorphism:
\[
  \re_{\ext_0} : \re\,(\Gamma \ext_0 A) \simeq ((\gamma : \re\,\Gamma)\times \Tm_{0,j}\,(\re\,A\,\gamma))
\]
This is because substitutions targeting $\Gamma \ext_0 A$ are the same as
pairs of substitutions and terms, by the specification of $\blank\ext_0\blank$.

\subsection{The Logical Relation}

Internally to $\hato$, we define by induction on the syntax of 2LTT a
proof-relevant logical relation interpretation, called $\blank^R$. The induction
motives are specified as follows.
\begin{alignat*}{5}
  &(\Gamma &&: \Con)^R && : \ev\,\Gamma \to \re\,\Gamma \to \whset_\omega\\
  &(\sigma &&: \Sub\,\Gamma\,\Delta)^R &&: \Gamma^R\,\gamma\,\gamma' \to \Delta^R\,(\ev\,\sigma\,\gamma)\,(\re\,\sigma\,\gamma')\\
  &(A      &&: \Ty_{1,j}\,\Gamma)^R &&: \Gamma^R\,\gamma\,\gamma' \to \ev\,A\,\gamma \to \re\,A\,\gamma' \to \whset_j\\
  &(t      &&: \Tm_{1,j}\,\Gamma\,A)^R &&: (\gamma^R : \Gamma^R\,\gamma\,\gamma') \to A^R\,\gamma^R\,(\ev\,t\,\gamma)\,(\re\,t\,\gamma')\\
  &(A      &&: \Ty_{0,j}\,\Gamma)^R &&: \Gamma^R\,\gamma\,\gamma' \to \emb{\ev\,A\,\gamma} = \re\,A\,\gamma' \\
  &(t      &&: \Tm_{0,j}\,\Gamma\,A)^R &&: \Gamma^R\,\gamma\,\gamma' \to \emb{\ev\,t\,\gamma} = \re\,t\,\gamma'
\end{alignat*}
For $\Con$, $\Sub$, and meta-level types and terms, this is a fairly standard
logical relation interpretation: contexts are mapped to relations, types to
dependent relations, and substitutions and terms respect relations. For
object-level types and terms, we get soundness statements: evaluation via $\ev$
followed by embedding back to 2LTT is the same as restricting 2LTT syntax to
object-level contexts.

\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

\end{document}
\endinput