### 2024-02-05 dept23prez/prez.tex
just push everything

\documentclass[dvipsnames,aspectratio=169]{beamer}
\usetheme{Madrid}

%% kill footline
\setbeamertemplate{footline}[frame number]{}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}{}

%% bibliography
\bibliographystyle{alpha}
\setbeamerfont{bibliography item}{size=\footnotesize}
\setbeamerfont{bibliography entry author}{size=\footnotesize}
\setbeamerfont{bibliography entry title}{size=\footnotesize}
\setbeamerfont{bibliography entry location}{size=\footnotesize}
\setbeamerfont{bibliography entry note}{size=\footnotesize}
\setbeamertemplate{bibliography item}{}

%% kill ball enumeration
\setbeamertemplate{enumerate items}[circle]
\setbeamertemplate{section in toc}[circle]

%% kill block shadows
\setbeamertemplate{blocks}[rounded][shadow=false]
\setbeamertemplate{title page}[default][colsep=-4bp,rounded=true]

%% kill ball itemize
\setbeamertemplate{itemize items}[circle]

%% --------------------------------------------------------------------------------

\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{cite}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{mathpartir}
\usepackage{scalerel}
\usepackage{stmaryrd}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{fontspec}

\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  urlcolor=blue}

\setmonofont[SizeFeatures={Size=8}]{DejaVu Sans Mono}

%% --------------------------------------------------------------------------------

\renewcommand{\mit}[1]{\mathit{#1}}
\newcommand{\msf}[1]{\mathsf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mbf}[1]{\mathbf{#1}}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ext}{\triangleright}
\newcommand{\Code}{\msf{Code}}
\newcommand{\El}{\msf{El}}
\newcommand{\lam}{\msf{lam}}
\newcommand{\app}{\msf{app}}
\newcommand{\NatElim}{\msf{NatElim}}
\newcommand{\y}{\msf{y}}

\newcommand{\Lift}{{\Uparrow}}
\newcommand{\spl}{{\sim}}
\newcommand{\qut}[1]{{<}#1{>}}

\newcommand{\mbbc}{\mbb{C}}
\newcommand{\mbbo}{\mbb{O}}
\newcommand{\ob}{_\mbbo}

\newcommand{\U}{\msf{U}}
\newcommand{\Con}{\msf{Con}}
\newcommand{\Sub}{\msf{Sub}}
\newcommand{\Ty}{\msf{Ty}}
\newcommand{\Tm}{\msf{Tm}}
\newcommand{\Cono}{\msf{Con}_{\mbbo}}
\newcommand{\Subo}{\msf{Sub}_{\mbbo}}
\newcommand{\Tyo}{\msf{Ty}_{\mbbo}}
\newcommand{\Tmo}{\msf{Tm}_{\mbbo}}
\newcommand{\hCon}{\wh{\msf{Con}}}
\newcommand{\hSub}{\wh{\msf{Sub}}}
\newcommand{\hTy}{\wh{\msf{Ty}}}
\newcommand{\hTm}{\wh{\msf{Tm}}}

\newcommand{\p}{\mathsf{p}}
\newcommand{\q}{\mathsf{q}}

\newcommand{\refl}{\msf{refl}}
\newcommand{\Bool}{\msf{Bool}}
\newcommand{\true}{\msf{true}}
\newcommand{\false}{\msf{false}}
\newcommand{\True}{\msf{True}}
\newcommand{\False}{\msf{False}}
\newcommand{\List}{\msf{List}}
\newcommand{\nil}{\msf{nil}}
\newcommand{\cons}{\msf{cons}}
\newcommand{\Nat}{\msf{Nat}}
\newcommand{\zero}{\msf{zero}}
\newcommand{\suc}{\msf{suc}}
\renewcommand{\tt}{\msf{tt}}
\newcommand{\fst}{\msf{fst}}
\newcommand{\snd}{\msf{snd}}
\newcommand{\mylet}{\msf{let}}
\newcommand{\emptycon}{\scaleobj{.75}\bullet}
\newcommand{\id}{\msf{id}}

\newcommand{\Set}{\mathsf{Set}}
\newcommand{\Prop}{\mathsf{Prop}}
\newcommand{\Rep}{\msf{Rep}}
\newcommand{\blank}{{\mathord{\hspace{1pt}\text{--}\hspace{1pt}}}}
\newcommand{\emb}[1]{\ulcorner#1\urcorner}

\newcommand{\Stage}{\msf{Stage}}
\newcommand{\hato}{\bm\hat{\mbbo}}
\newcommand{\ev}{\mbb{E}}
\newcommand{\re}{\mbb{R}}

\theoremstyle{remark}
\newtheorem{notation}{Notation}

\newcommand{\whset}{\wh{\Set}}
\newcommand{\rexti}{\re_{\ext_1}^{-1}}
\newcommand{\rextizero}{\re_{\ext_0}^{-1}}

\newcommand{\rel}{^{\approx}}
\newcommand{\yon}{\msf{y}}

\newcommand{\Vect}{\msf{Vec}}
\newcommand{\msA}{\msf{A}}
\newcommand{\msB}{\msf{B}}
\newcommand{\mst}{\msf{t}}
\newcommand{\msu}{\msf{u}}
\newcommand{\msv}{\msf{v}}
\newcommand{\msx}{\msf{x}}
\newcommand{\msy}{\msf{y}}
\newcommand{\msz}{\msf{z}}
\newcommand{\msas}{\msf{as}}
\newcommand{\msbs}{\msf{bs}}
\newcommand{\msxs}{\msf{xs}}
\newcommand{\msys}{\msf{ys}}
\newcommand{\msff}{\msf{f}}
\newcommand{\msg}{\msf{g}}
\newcommand{\msh}{\msf{h}}
\newcommand{\msa}{\msf{a}}
\newcommand{\msn}{\msf{n}}

%% --------------------------------------------------------------------------------

\title{Dysfunctional Programming}
\author{András Kovács}
\institute{
  {Eötvös Loránd University, Dept. of Programming Languages and Compilers}
}
\date{9 February 2023, Department Workshop}
\begin{document}

\frame{\titlepage}

%% ------------------------------------------------------------

\begin{frame}{First-class functions}

A huge amount of software abstraction is based on first-class functions:
\begin{itemize}
\item HOFs in FP
\item Virtual methods in OOP
\item Callbacks, closures in many settings
\end{itemize}
\vspace{1em}
\pause

Lots of headache in compilation:
\begin{itemize}
\item Dynamic calls block static analysis \& inlining
\item Implementation of closures/vtables adds complexity
\end{itemize}
\vspace{1em}
\pause

\begin{block}{}
\textbf{What can we do without first-class functions?}
\end{block}

\end{frame}

%% ------------------------------------------------------------

\begin{frame}[fragile]{}

Do we need function values here?

\begin{exampleblock}{}
\begin{verbatim}
    map : (A → B) → List A → List B
    map f nil         = nil
    map f (cons a as) = cons (f a) (map f as)

    g : List Int → List Int
    g = map (+1)
\end{verbatim}
\end{exampleblock}
\pause

Instead:

\begin{exampleblock}{}
\begin{verbatim}
    g : List Int → List Int
    g nil         = nil
    g (cons a as) = cons (a + 1) (g as)
\end{verbatim}
\end{exampleblock}
\pause
Function elimination by inlining.

\end{frame}

%% ------------------------------------------------------------

\begin{frame}{Getting rid of function values}

  \begin{enumerate}
  \item Metaprogramming: use first-class functions only at compile time, generate first-order code.
        Inlining is a special case of this.
  \item Defunctionalization: replace closures by explicit first-order data.
  \end{enumerate}
\vspace{1em}

We focus on \textbf{1}, using \textbf{two-level type theory}.
\vspace{1em}
\pause

There is a dependently typed compile-time language for metaprogramming.
\vspace{1em}
\pause

Object-level \& meta-level code is freely mixed in programs.
\vspace{1em}
\pause

Generated code is guaranteed to be \emph{well-typed}, \emph{simply typed} and \emph{only have first-order functions}.

\end{frame}

%% ------------------------------------------------------------

\begin{frame}[fragile]{}
\begin{block}{Rules, informally}
  \begin{itemize}
  \item Compile-time types and runtime types have different types (``universes'').
  \item Function and non-function (``value'') types at runtime stage are distinguished.
  \item There are only first-order function types at runtime.
  \item We can't store functions in runtime data.
  \end{itemize}
\end{block}

\begin{exampleblock}{}
\begin{verbatim}
    data List (A : Ty Val) : Ty Val where
      nil  : List A
      cons : A → List A → List A

    map : {A B : Ty Val} → (A → B) → List A → List B

    g : List Int -> List Int
    g := map (+1)
\end{verbatim}
\end{exampleblock}
\end{frame}

%% ------------------------------------------------------------

\begin{frame}[fragile]{Is List a Functor/Applicative/Monad?}

OK:
\begin{exampleblock}{}
\begin{verbatim}
    class Functor (F : Ty Val → Ty Val) where
      fmap : (A → B) → F A → F B
\end{verbatim}
\end{exampleblock}
\pause

Issue with \texttt{Applicative}:
\begin{exampleblock}{}
\begin{verbatim}
    class Applicative (F : Ty Val → Ty Val) where
      pure : A → F A
      ap   : F (A → B) → F A → F B
\end{verbatim}
\end{exampleblock}
\pause

Solution:

\begin{exampleblock}{}
\begin{verbatim}
    class Applicative (F : Ty Val → Ty Val) where
      point : F ()
      zip   : F A → F B → F (A, B)
\end{verbatim}
\end{exampleblock}
\pause

Monad is also OK:

\begin{exampleblock}{}
\begin{verbatim}
    bind : M A → (A → M B) → M B
\end{verbatim}
\end{exampleblock}
\end{frame}

\begin{frame}[fragile]{Is State a Functor/Applicative/Monad?}

\begin{exampleblock}{}
\begin{verbatim}
    State : Ty Val → Ty Val → MetaTy
    State S A = S → (A, S)
\end{verbatim}
\end{exampleblock}
\vspace{1em}

Using the previous Functor/Applicative/Monad definitions, no!
\vspace{1em}
\pause

But it is a \emph{relative} Functor/Applicative/Monad:

\begin{exampleblock}{}
\begin{verbatim}
    class RelFunctor (F : Ty Val → MetaTy) where ...
    class RelApplicative (F : Ty Val → MetaTy) where ...
    class RelMonad (M : Ty Val → MetaTy) where ...
    instance RelMonad (State S) where ...
\end{verbatim}
\end{exampleblock}
\end{frame}

\begin{frame}[fragile]{Defunctionalization}

Difference lists in Haskell:
\vspace{1em}

\begin{exampleblock}{}
\begin{verbatim}
    type DList a = [a] -> [a]

    append :: DList a -> DList a -> DList a
    append xs ys = xs . ys

    toDList a as = a : as
    fromDList xs = xs []
\end{verbatim}
\end{exampleblock}
\end{frame}

\begin{frame}[fragile]{Defunctionalization}

No functions needed:
\vspace{1em}

\begin{exampleblock}{}
\begin{verbatim}
    data DList a = Chunk [a] | Append (DList a) (DList a)

    apply :: DList a -> [a] -> [a]
    apply (Chunk xs)     ys = xs ++ ys
    apply (Append xs ys) zs = apply xs (apply ys zs)

    toDList :: [a] -> DList a
    toDList xs = Chunk xs

    fromDList :: DList a -> [a]
    fromDList xs = apply xs []
\end{verbatim}
\end{exampleblock}
\end{frame}

\begin{frame}{Summary \& further topics}

\begin{itemize}
\item Not having first-class functions makes compilation much easier.
  \begin{itemize}
  \item In most cases: metaprogramming can replace first-class functions.
  \item In most remaining cases: defunctionalization can replace first-class functions.
  \item In the remaining cases: we can distinguish closures \emph{in the type system}.
  \end{itemize}
  \vspace{1em}
\item Trade-off between programming convenience and performance.
  \vspace{1em}
\item Trade-off between code size and performance.
  \vspace{1em}
\item Other things to try to do without functions:
  \begin{itemize}
  \item Fusion optimizations.
  \item More things from the Haskell/ML literature.
  \end{itemize}
\end{itemize}

\end{frame}
\end{document}